---
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
---

# RNAseq downstream analysis      

```{r message=FALSE, warning=FALSE, include=FALSE}
#necessary packages and functions  
source("I:/Shared_win/projects/RNA_normal/analysis.r")
```


## load expression mtx       

```{r echo=FALSE}
# specify matrix path and name to save files then
dat = '../output/RNAseq..counts.gene.matrix'
mtxname = 'SS2_HC_20210812'

# specify names of 'A vs B', and cell idx of As and Bs
Aname = "LI_CKO_INF" 
Bname = "LI_CTL_INF"  

Aidx = c(1:4)
Bidx = c(5:7)

# specify filtering cutoffs
cut_count = 2 
cut_cell = 3

#
mat_raw <- read.table(dat, header = TRUE, stringsAsFactors = F , sep = "\t")
rownames(mat_raw) <- mat_raw$gene_id
mat_raw <- mat_raw[,2:ncol(mat_raw)]

mat_raw <- edgeR::cpm(mat_raw)

# round the reads count from rsem output
mat_raw <- round(mat_raw)

# choose only protein_coding genes
list_pc <- 'I:/Shared_win/genomics/mouse/GRCm38_vM25/gtf_detail/list_pc.lv1_2'
id_pc <- as.vector(unlist(read.table(list_pc)))
mat_pc <- mat_raw[id_pc,]

# filter for genes that expressed (CPM > 2) in at least 3 cells
#mat_pc <- mat_pc[rowSums(mat_pc[,c(Aidx,Bidx)] >cut_count) >= cut_cell,]
mat_pc <- mat_pc[rowSums(mat_pc[,c(Aidx)] >cut_count) >= cut_cell |
                   rowSums(mat_pc[,c(Bidx)] >cut_count) >= cut_cell,]
#matz_pc <- zscore_mat(log2(mat_pc+1))
```




check raw counts matrix       
```{r}
dim(mat_raw)
head(mat_raw)
```
       
check filtered pc(protein coding) matrix        
```{r}
dim(mat_pc[,c(Aidx,Bidx)])
head(mat_pc[,c(Aidx,Bidx)])
``` 
 
```{r echo=FALSE}
cat(paste0("get ",dim(mat_pc)[1]," filtered (cpm > ",cut_count," in at least ",cut_cell," cells) protein-coding genes for next analysis"))
```

```{r eval=FALSE, include=FALSE}
# save CPM matrix
mat_s <- mat_raw
write.table(cbind.data.frame(gene=rownames(mat_s),mat_s),paste0("RNAseq.",mtxname,".raw_cpm.gene.csv"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = ",")
mat_s1 <- mat_pc
write.table(cbind.data.frame(gene=rownames(mat_s1),mat_s1),paste0("RNAseq.",mtxname,".filt_cpm.pc_gene.csv"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = ",")

# save TPM matrix
datt = '../output/RNAseq.20210728_SS2_WYS.tpm.gene.matrix'
matt_raw <- read.table(datt, header = TRUE, stringsAsFactors = F , sep = "\t")
rownames(matt_raw) <- matt_raw$gene_id
matt_raw <- matt_raw[,2:ncol(matt_raw)]

write.table(cbind.data.frame(gene=rownames(matt_raw),matt_raw),paste0("RNAseq.",mtxname,".raw_tpm.gene.csv"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = ",")

matt_pc <- matt_raw[rownames(mat_pc),]
mat_s2 <- matt_pc
write.table(cbind.data.frame(gene=rownames(mat_s2),mat_s2),paste0("RNAseq.",mtxname,".filt_tpm.pc_gene.csv"),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = ",")
```

## QC  

design info  
```{r echo=FALSE, paged.print=FALSE}
design <- data.frame(condition=factor(c(rep(Aname,length(Aidx)),rep(Bname,length(Bidx)))),
#                     replicate=factor(c(paste0("rep",1:length(Aidx)),paste0("rep",1:length(Bidx)))))
                     replicate=factor(colnames(mat_pc[,c(Aidx,Bidx)])))
rownames(design) <- colnames(mat_pc[,c(Aidx,Bidx)])

design
```


### condition mean expression              

```{r cross1, echo=FALSE, fig.height=6, fig.width=6, warning=FALSE}
plot(data.frame(Aname=rowMeans(mat_raw[rowSums(mat_raw[,c(Aidx,Bidx)])!=0,][,Aidx]),
                Bname=rowMeans(mat_raw[rowSums(mat_raw[,c(Aidx,Bidx)])!=0,][,Bidx])), pch=20, log = 'xy',
     xlab = paste0("Mean normalized counts (",Aname,")"),xlim=c(0.1,100000),
     ylab = paste0("Mean normalized counts (",Bname,")"),ylim=c(0.1,100000),
     main = paste0("expressed ",dim(mat_raw[rowSums(mat_raw)!=0,])[1]," genes of total ",dim(mat_raw)[1]," genes "))
```


```{r cross2, echo=FALSE, fig.height=6, fig.width=6, warning=FALSE}
plot(data.frame(Aname=rowMeans(mat_pc[,Aidx]),
                Bname=rowMeans(mat_pc[,Bidx])), pch=20, log = 'xy',
     xlab = paste0("Mean normalized counts (",Aname,")"),xlim=c(0.1,100000),
     ylab = paste0("Mean normalized counts (",Bname,")"),ylim=c(0.1,100000),
     main = paste0(dim(mat_pc)[1]," filtered protein coding genes "))
```


### PCA          

```{r include=FALSE}
rv <- rowVars(mat_pc[,c(Aidx,Bidx)])
selt <- order(rv, decreasing = TRUE)[seq_len(2000)]
pca2 <- stats::prcomp(t(mat_pc[selt,c(Aidx,Bidx)]), scale.=TRUE, center= TRUE)
#pca2$x

pca_d <- as.data.frame(pca2$x)

pca_d[,"condition"] = design$condition
pca_d[,"batch"] = rownames(design)
pca_d[,"replicate"] = design$replicate
```


```{r echo=FALSE}
plot(pca2$sdev,pch=20, xlab="PC",ylab="Standard Variation", main="PCs of top2000 highly variable genes")
```

##### summary of PCs
```{r echo=FALSE}
summary(pca2)
```



```{r echo=FALSE, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
#plot the PCA
#create a classic 2-dimension PCA plot (first two principal components) with conditions and library methods indicated
cols <- c("#481567FF","#1F968BFF","#B5640B")
#cols <- c("Gut" = cols[1], "Lung" = cols[2], "Meninges"=cols[3])
cols <- c(cols[3], cols[2])
names(cols) <- c(as.character(unique(design$condition)[1]),as.character(unique(design$condition)[2]))

#p1 = ggplot(data=pca_d, aes(x=PC1, y=PC2, color=condition, shape=condition))
p1 = ggplot(data=pca_d, aes(x=PC1, y=PC2, color=condition))
p1 = p1 + geom_point(size=3.5)
p1 = p1 + ggrepel::geom_text_repel(mapping = aes(label=pca_d$batch),size=2.5)
#p1 = p1 + stat_ellipse(type="norm", linetype=2)
p1 = p1 + labs(title="PCA\nusing CPM(top2000 highly variable genes) ",
               color="Condition", shape="batch")
p1 = p1 + scale_colour_manual(values = cols, limits=rev) 
#p1 = p1 + scale_shape_manual(values=c(16,17,15))
p1
```


```{r echo=FALSE, fig.height=3, fig.width=4.2, message=FALSE, warning=FALSE}
#plot the PCA
#create a classic 2-dimension PCA plot (first two principal components) with conditions and library methods indicated
cols <- c("#481567FF","#1F968BFF","#B5640B")
#cols <- c("Gut" = cols[1], "Lung" = cols[2], "Meninges"=cols[3])
cols <- c(cols[3], cols[2])
names(cols) <- c(as.character(unique(design$condition)[1]),as.character(unique(design$condition)[2]))

#p1 = ggplot(data=pca_d, aes(x=PC1, y=PC2, color=condition, shape=condition))
p1 = ggplot(data=pca_d, aes(x=PC1, y=PC2, color=condition))
p1 = p1 + geom_point(size=3.5)
#p1 = p1 + ggrepel::geom_text_repel(mapping = aes(label=pca_d$batch),size=2.5)
#p1 = p1 + stat_ellipse(type="norm", linetype=2)
p1 = p1 + labs(title="PCA",
               color="Condition", shape="batch")
p1 = p1 + scale_colour_manual(values = cols, limits=rev) + theme_classic()
#p1 = p1 + scale_shape_manual(values=c(16,17,15))
p1
```



```{r eval=FALSE, include=FALSE}
ggsave("./figures/LI_cnt2.CKO_INF_vs_CTL_INF/PCA.pdf",
       plot = p1,
       width = 4.5, height = 3)
```


```{r echo=FALSE, fig.height=4.5, fig.width=6, message=FALSE, warning=FALSE}
#plot the PCA
#create a classic 2-dimension PCA plot (first two principal components) with conditions and library methods indicated
#cols <- c("#481567FF","#1F968BFF","#B5640B")
#cols <- c("Gut" = cols[1], "Lung" = cols[2], "Meninges"=cols[3])
#cols <- c( "Meninges"=cols[3],"Lung" = cols[2])
#p2 = ggplot(data=pca_d, aes(x=PC1, y=PC2, color=condition, shape=condition))
p2 = ggplot(data=pca_d, aes(x=PC3, y=PC4, color=condition))
p2 = p2 + geom_point(size=3.5)
p2 = p2 + ggrepel::geom_text_repel(mapping = aes(label=pca_d$batch),size=2.5)
#p2 = p2 + stat_ellipse(type="norm", linetype=2)
p2 = p2 + labs(title="PCA\nusing CPM(top2000 highly variable genes) ",
               color="Condition", shape="batch")
p2 = p2 + scale_colour_manual(values = cols) 
#p2 = p2 + scale_shape_manual(values=c(16,17,15))
p2
```


### tSNE       

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=8,fig.height=6}
set.seed(131)
PCs <- 1:6
PCss <- paste0(PCs[1],":",PCs[length(PCs)])
pca2.emb <- pca2$x %*% diag(pca2$sdev**2)
tsne_out <- Rtsne::Rtsne(
  as.matrix(pca2.emb[,PCs]),
  pca = FALSE,
  perplexity = 2,
  max_iter = 2000
)

#cols <- c("Meninges" = "#481567FF", "nonMeninges" = "#1F968BFF")
p6 = ggplot(data=data.frame(tsne_out$Y), aes(x=X1, y=X2)) + labs(x="tSNE_1",y="tSNE_2", seed=FALSE) + 
  #geom_point(aes(colour=design$condition, shape=design$condition),size=3.5) + 
  geom_point(aes(colour=design$condition),size=3.5) + 
  #stat_ellipse(aes(colour=design$condition),type="norm", linetype=2) + 
   ggrepel::geom_text_repel(mapping = aes(label=colnames(mat_pc[,c(Aidx,Bidx)])),size=2.5) + 
  scale_colour_manual(values = cols) + 
  #scale_shape_manual(values=c(16:10)) + 
   labs(title=paste0("tSNE\nusing PC",PCss),
               color="Condition", shape="batch")
p6
```



```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=4.2,fig.height=3}
set.seed(131)
PCs <- 1:6
PCss <- paste0(PCs[1],":",PCs[length(PCs)])
pca2.emb <- pca2$x %*% diag(pca2$sdev**2)
tsne_out <- Rtsne::Rtsne(
  as.matrix(pca2.emb[,PCs]),
  pca = FALSE,
  perplexity = 2,
  max_iter = 2000
)

#cols <- c("Meninges" = "#481567FF", "nonMeninges" = "#1F968BFF")
p6.1 = ggplot(data=data.frame(tsne_out$Y), aes(x=X1, y=X2)) + labs(x="tSNE_1",y="tSNE_2", seed=FALSE) + 
  #geom_point(aes(colour=design$condition, shape=design$condition),size=3.5) + 
  geom_point(aes(colour=design$condition),size=3.5) + 
  #stat_ellipse(aes(colour=design$condition),type="norm", linetype=2) + 
   #ggrepel::geom_text_repel(mapping = aes(label=colnames(mat_pc[,c(Aidx,Bidx)])),size=2.5) + 
  scale_colour_manual(values = cols, limits=rev) + 
  #scale_shape_manual(values=c(16:10)) + 
   labs(title="tSNE",
               color="Condition", shape="batch") + theme_classic()
p6.1
```


```{r eval=FALSE, include=FALSE}
ggsave("./figures/LI_cnt2.CKO_INF_vs_CTL_INF/tSNE.pdf",
       plot = p6.1,
       width = 4.5, height = 3)
```

### correlation        

```{r include=FALSE}
#define a function to plot correlation  
plotCor <- function(Mat){
  library(edgeR)
  library(corrplot)
  Cor <- cor(log2(edgeR::cpm(Mat)+1))
  par(cex=0.54, pin=c(8,8))
  corrplot(Cor,method="number",title = "pearson correlation of log2(CPM+1)",mar = c(0, 0, 1, 0))
}
```



correlation (log2(CPM+1)) of all filtered protein-coding genes  
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6.8, fig.height=6.8}
plotCor(mat_pc)
```

correlation (log2(CPM+1)) of all filtered protein-coding genes, grouping by correlation  

```{r echo=FALSE,message=FALSE,warning=FALSE,fig.width=5,fig.height=5}
  Ret_mat <- log2(mat_pc[,c(Aidx,Bidx)]+1)
  Ret_cor <- cor(Ret_mat)
  
  lower = 0.92
  upper = 1
  pal <- "Reds"
  
  ht1 <- ComplexHeatmap::Heatmap(Ret_cor, col = circlize::colorRamp2(seq(lower, upper, ((upper-lower)/7)),RColorBrewer::brewer.pal(8, pal)),
                  heatmap_legend_param = list(
                  color_bar = "continuous",
                  legend_direction = "horizontal",
                  legend_width = unit(5, "cm"),
                  title_position = "topcenter"),
                  name = "Pearson correlation",
                  column_names_gp = grid::gpar(fontsize = 10),
                  row_names_gp = grid::gpar(fontsize = 10),
                  top_annotation = NULL)
  heatplot <- ComplexHeatmap::draw(ht1, heatmap_legend_side = "top")
``` 


## Differential Analysis                


```{r echo=FALSE}
ret <- run_MEedgeR(MAT = mat_pc[,c(Aidx,Bidx)], ED = Aname, WT = Bname, n1 = length(Aidx),n2 = length(Bidx),
                         lcm = 0, padj = 0.05,lfc = log2(1.5), gsea = TRUE)
```


```{r}
ret$SUM
```

```{r eval=FALSE, include=FALSE, paged.print=FALSE}
ret$gsea
```


##### using cutoff: *p.value* 0.05, FC 1.5     

```{r echo=FALSE, message=FALSE, warning=FALSE}
rets <- finalplot(mat_pc[,c(Aidx,Bidx)],data.frame(ret$DEG_edgeR), paste0(Aname,"_vs_",Bname), 0.05, 1.5, 
                  Sign = FALSE, Sign_dn = 40, Sign_up = 40, padjust = F)
```



```{r echo=FALSE, fig.height=8, fig.width=12, warning=FALSE}
rets$vol
```
 
```{r echo=FALSE, fig.height=8, fig.width=12, warning=FALSE}
rets$vol
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
rets2 <- finalplot(mat_pc[,c(Aidx,Bidx)],data.frame(ret$DEG_edgeR), paste0(Aname,"_vs_",Bname), 0.05, 1.5, 
                  Sign = FALSE, Sign_dn = 40, Sign_up = 40, padjust = F, Label = "")
```

```{r echo=FALSE, fig.height=2.5, fig.width=4.6, warning=FALSE}
rets2$vol
```

```{r eval=FALSE, include=FALSE}
ggsave("./figures/LI_cnt2.CKO_INF_vs_CTL_INF/volcano.pvalue.pdf",
       plot = rets2$vol,
       width = 5, height = 2.5)
``` 


```{r echo=FALSE, message=FALSE, warning=FALSE}
rets1 <- finalplot(mat_pc[,c(Aidx,Bidx)],data.frame(ret$DEG_edgeR), paste0(Aname,"_vs_",Bname), 0.05, 1.5, 
                  Sign = FALSE, Sign_dn = 40, Sign_up = 40, padjust = TRUE, Label = "")
```

```{r echo=FALSE, fig.height=2.5, fig.width=4.6, warning=FALSE}
rets1$vol
``` 
 
 

```{r eval=FALSE, include=FALSE}
ggsave("./figures/LI_cnt2.CKO_INF_vs_CTL_INF/volcano.pdf",
       plot = rets1$vol,
       width = 5, height = 2.5)
``` 
 
```{r echo=FALSE, fig.height=9.6, fig.width=4.8}
genes <- rets$marker
pheatmap::pheatmap(log2(mat_pc[genes,c(Aidx,Bidx)]+1),cluster_rows = F, cluster_cols = F,
                   fontsize_row = 8,
         gaps_col = length(Aidx), 
         gaps_row = 40,
         main = "top log2(CPM+1)")

```
 
```{r echo=FALSE, fig.height=9.6, fig.width=7.2}
pheatmap(log2(mat_raw[genes,]+1),cluster_rows = F, cluster_cols = F,
         gaps_col = c(4,7,11,15,18),
         gaps_row = 40,
         main = "top log2(CPM+1)")

```
 
```{r echo=FALSE, fig.height=9.6, fig.width=4.8}
pheatmap(zscore_mat(log2(mat_pc[genes,c(Aidx,Bidx)]+1)),cluster_rows = F, cluster_cols = F,
         fontsize_row = 8,
         gaps_col = length(Aidx),
         gaps_row = 40,
         main = "top zscore")

```
 
```{r echo=FALSE, fig.height=9.6, fig.width=7.2}
pheatmap(zscore_mat(log2(mat_pc[genes,]+1)),cluster_rows = F, cluster_cols = F,
         fontsize_row = 8,
         gaps_col = c(4,7,11,15,18),
         gaps_row = 40,
         main = "top zscore")

```
 
```{r eval=FALSE, include=FALSE}
write.table(rec_gcol(data.frame(ret$DEG_edgeR)), paste0("edgeR_DEGs.",
           #paste(colnames(mat_pc)[Aidx],collapse="."),"_vs_", paste(colnames(mat_pc)[Bidx],collapse="."),".csv"),
           paste(Aname),"_vs_", paste(Bname),".csv"),
           col.names = TRUE, row.names = FALSE,quote=FALSE, sep=",")
```
 
```{r eval=FALSE, include=FALSE}
proc.mat <- function(Mat){
  Mat <- cbind(rownames(Mat),"NA",Mat)
  colnames(Mat)[1:2] <- c("NAME","DESCRIPTION")
  return(Mat)
}

write.table(proc.mat(Mat=ret$TMM$counts[rownames(ret$DEG_edgeR),]),
            paste0("GSEA_mit/RNAseq.",mtxname,".",Aname,"_vs_",Bname,".filt_tmm.pc_gene.txt"),
            col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)

write.table(c(paste0(length(c(Aidx,Bidx))," 2 1"),
              paste0("# ",Aname," ",Bname),
              paste(c(rep("1",length(Aidx)),rep("0",length(Bidx))),collapse = " ")),
            paste0("GSEA_mit/RNAseq.",mtxname,".",Aname,"_vs_",Bname,".cls"),
            col.names = FALSE, row.names = FALSE, sep = "\t", quote = FALSE)
```
 

 

 
```{r echo=FALSE, fig.height=3.6, fig.width=4}
#genes <- rets$marker
genes <- c(rets$down[1],rev(rets$up[1:7]))

pheatmap::pheatmap(log2(mat_pc[genes,c(rev(Bidx),Aidx)]+1),cluster_rows = F, cluster_cols = F,
                   fontsize_row =8,
         gaps_col = length(Bidx), border_color = NA,
         gaps_row = 1,
         main = "top log2(CPM+1)")

```

 
```{r echo=FALSE, fig.height=3.6, fig.width=4}
pheatmap::pheatmap(zscore_mat(log2(mat_pc[genes,c(rev(Bidx),Aidx)]+1)),cluster_rows = F, cluster_cols = F,
         fontsize_row =8,
         gaps_col = length(Bidx),
         gaps_row =1, color = color.test,border_color = NA,
         breaks = seq(-2,2,0.04),
         main = "top zscore")

```
 
```{r eval=FALSE, include=FALSE}
pdf("./figures/LI_cnt2.CKO_INF_vs_CTL_INF/heat1.pdf",
    width = 4, height = 3.6)
pheatmap::pheatmap(log2(mat_pc[genes,c(rev(Bidx),Aidx)]+1),cluster_rows = F, cluster_cols = F,
                   fontsize_row =8,
         gaps_col = length(Bidx), border_color = NA,
         gaps_row = 1,
         main = "top log2(CPM+1)")

dev.off()
pdf("./figures/LI_cnt2.CKO_INF_vs_CTL_INF/heat1z.pdf",
    width = 4, height = 3.6)
pheatmap::pheatmap(zscore_mat(log2(mat_pc[genes,c(rev(Bidx),Aidx)]+1)),cluster_rows = F, cluster_cols = F,
         fontsize_row =8,
         gaps_col = length(Bidx),
         gaps_row =1, color = color.test,border_color = NA,
         breaks = seq(-2,2,0.04),
         main = "top zscore")
dev.off()

#

```
 
 
 
 
```{r echo=FALSE, fig.height=9.6, fig.width=4}
#genes <- rets$marker
genes <- c(rets$down[1:40],rev(rets$up[1:40]))

pheatmap::pheatmap(log2(mat_pc[genes,c(rev(Bidx),Aidx)]+1),cluster_rows = F, cluster_cols = F,
                   fontsize_row =7.6,
         gaps_col = length(Bidx), border_color = NA,
         gaps_row = 40,
         main = "top log2(CPM+1)")

```

 
```{r echo=FALSE, fig.height=9.6, fig.width=4}
pheatmap::pheatmap(zscore_mat(log2(mat_pc[genes,c(rev(Bidx),Aidx)]+1)),cluster_rows = F, cluster_cols = F,
         fontsize_row =7.6,
         gaps_col = length(Bidx),
         gaps_row = 40, color = color.test,border_color = NA,
         breaks = seq(-2,2,0.04),
         main = "top zscore")

```
 
 
 
```{r eval=FALSE, include=FALSE}
pdf("./figures/LI_cnt2.CKO_INF_vs_CTL_INF/heat2.pdf",
    width = 4, height = 9.6)
pheatmap::pheatmap(log2(mat_pc[genes,c(rev(Bidx),Aidx)]+1),cluster_rows = F, cluster_cols = F,
                   fontsize_row =7.6,
         gaps_col = length(Bidx), border_color = NA,
         gaps_row = 40,
         main = "top log2(CPM+1)")

dev.off()
pdf("./figures/LI_cnt2.CKO_INF_vs_CTL_INF/heat2z.pdf",
    width = 4, height = 9.6)
pheatmap::pheatmap(zscore_mat(log2(mat_pc[genes,c(rev(Bidx),Aidx)]+1)),cluster_rows = F, cluster_cols = F,
         fontsize_row =7.6,
         gaps_col = length(Bidx),
         gaps_row = 40, color = color.test,border_color = NA,
         breaks = seq(-2,2,0.04),
         main = "top zscore")
dev.off()

#

```
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 